Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2023-02-14 paper/paper.tex
Inline verification conditions

\section{\agr}
\label{sec:agree}
\input{verification-conditions}

>>>>

\section{\agr}
\label{sec:agree}

\newcommand{\globally}{\konst{Always}}
\newcommand{\historically}{\konst{Hist}}
\newcommand{\assumes}{\ensuremath{A}}
\newcommand{\guarantees}{\ensuremath{P}}
\newcommand{\inputs}{\ensuremath{I}}
\newcommand{\outputs}{\ensuremath{O}}
\newcommand{\components}{\ensuremath{C}}
\newcommand{\component}{\ensuremath{c}}

The \agr\ specification language, and semantics, are inspired by the Lustre language \cite{10.1145/41625.41641}.
%
Lustre semantics are synchronous data-flow where the inputs and outputs of components are streams.
%
Informally, streams assign values to expressions at each point in time.
%
\agr\ contracts express relationships between input and output streams by assuming a set of allowed input streams and then guaranteeing a corresponding set of allowed output streams.

%%
% COMPONENTS
%%

\subsection{Components}

A system implementation is described by its ports, its subcomponents, and the connections between ports on different subcomponents.
%
Ports in \agr\ are inherited from AADL.
%
There are three kinds of port: \emph{event} ports, \emph{data} ports, and \emph{event data} ports.
%
An event port carries no data and only signals occurrence (or not) of the event.
%
A data port is a sampled port: the data can be read at anytime.
%
An event data port indicates when data is present with an event.

Data flows through the subcomponents of a system in dependency order, with inputs being propagated to outputs through all contracts until they stabilize \ie, cannot propagate further.
%
Therefore, subcomponent contracts, and thus the top-level model, must be acyclic.
%
Cyclic systems are made acyclic by breaking cycles with delays.
%
Once the data propagation has stabilized, the model proceeds to the next input data in the input streams.
%
It is worth noting that the semantics do not model computation or communication delay: the output of one contract is seen at the input of any downstream contract in the same step of the input data stream.

%%
% CONTRACTS AND VERIFICATION CONDITIONS
%%

\subsection{Contracts and verification conditions}

From the system and component contracts, \agr\ generates a set of verification conditions to show that a system's component implementation is correct~\cite{agree2013}.
%
The \agr\ model checker is then invoked to prove or disprove the verification conditions.
%
The contracts and verification conditions are expressed in \emph{past-time linear temporal logic} (PLTL) \cite{10.1093/jigpal/8.1.55}.

PLTL is a logic enhanced with temporal operators able to reason about the truth values of formulas through time.
%
It combines past-time operators that reason over the history of computation up to a point (PTLTL) and future-time operators that reason over the computation that follows that point (LTL).
%
Its semantics are defined relative to a point in time $i$, a finite trace of system states $\pi = s_0, s_1, \ldots, s_i$, and some logic formula $f$.

The two PLTL operators necessary for the \agr\ generated verification conditions are $\globally$ that looks forward in time along the trace and $\historically$ (historically) that looks backward in time along the trace.
%
These are defined as
%
\begin{eqnarray*}
 (\pi, i) \models \globally(f) & \iff & \forall j \ge i, (\pi, j) \models f \\
(\pi, i) \models \historically(f) & \iff & \forall 0 \le j \le i, (\pi, j) \models f
\end{eqnarray*}
%
The $\models$-operator is read as \emph{satisfies}.
%
A trace at a moment in time satisfies $\globally(f)$ if and only if it satisfies $f$ in the current and all future states of $\pi$.
%
$\globally(f)$ is invariant from the current moment into the future and $\historically$ is invariant from the beginning of the trace to the current moment.

A \emph{system} is a tuple $(\inputs, \outputs, \assumes, \guarantees,C)$, where $\inputs$ is the input set, $\outputs$ is the output set, $\assumes$ is the set of assumptions, $\guarantees$ is the set of guarantees, and $C$ are subcomponents.
%
Assumptions and guarantees only use the past-time operators from PLTL.
%
A subcomponent $\component$ is, hierarchically, also a system, and may be designated by its own tuple $(\inputs_\component, \outputs_\component, \assumes_\component, \guarantees_\component, C_c)$.
%
From the components and their connections, $\mathbb{I}_\component$ is defined to be the set of components providing input to some component $\component$ in the system, and $\mathbb{O}$ is defined to be the set of components that provide the output for the system.

A system is \emph{correct} if and only if for all components $c \in C$ the following two verification conditions hold:
%
\begin{equation}\label{eq:assumes}
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{I}_\component} P_{\component^\prime})
            \implies \assumes_\component)
\end{equation}
%
\begin{equation}\label{eq:guarantees}
            \globally(\historically(\assumes \wedge
            \bigwedge_{\component^\prime \in \mathbb{O}} \guarantees_{\component^\prime})
            \implies \guarantees)
\end{equation}
%
The conditions from \eqref{eq:assumes} verify the input assumptions on each component under the system assumptions and upstream component guarantees.
%
They check if the component guarantees and system assumptions are strong enough to imply input assumptions on all immediate downstream components.
%
The condition in \eqref{eq:guarantees} checks the output guarantees of the system under the system assumptions and component guarantees that provide the output.
%
It checks if the guarantees on components providing primary outputs are strong enough to imply the system guarantees.

If all the verification conditions hold (\agr\ uses $k$-inductive model checking to automatically prove or disprove each generated verification condition), then the system is said to be \emph{correct}, meaning that the system composition meets input assumptions at each input as well as the guarantees on the system output.
%
A consequence of this result is that $\globally(\historically(\assumes) \implies \guarantees)$ holds for the system contract.
%
Therefore the Liskov substitution principle applies and the implementation is a safe substitution for the system contract \cite{10.1145/62139.62141}.

The expanded property lists in \figref{fig:example-certificate} and \figref{fig:hardened-certificate} are the results from verifying or disproving the above verification conditions.
%
The additional unexpanded results at the bottom of the figures prove \emph{self-consistency} in the contracts.
%
A contract is self-consistent if it does not guarantee two different values in the same moment of time, \ie, the contract is not self-contradicting.
%
\agr\ generates additional verification conditions that prove each component contract, and the composition of contracts, self-consistent.

%%
% SYNTAX
%%
